---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# formatstat

<!-- badges: start -->

[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

The goal of formatstat is to provide functions that format statistical
output in a way that can be inserted into R Markdown documents. This is
analogous to the
[`apa_print()`](https://frederikaust.com/papaja_man/reporting.html#statistical-models-and-tests)
functions in the [papaja](https://github.com/crsh/papaja) package but
can print Markdown or LaTeX syntax. The defaults follow [American
Psychological Association style](https://apastyle.apa.org/), but some
defaults can be over-ridden.

## Installation

You can install the development version of formatstat from
[GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("JeffreyRStevens/formatstat")
```

## Example

For an example, we'll create a data set from the `beavers1` and
`beavers2` data sets.

```{r example}
library(formatstat)
beavers <- merge(beaver1, beaver2, by = "time")
beavers <- beavers[, c("time", "temp.x", "temp.y")]
beavers_corr <- cor.test(beavers$temp.x, beavers$temp.y)
```

Now we can apply the `apa_ttest()` to `beavers_corr` to create a
Markdown- formatted character string for the statistical results. We can
embed this as inline R Markdown code to generate the results.

#### Code

`` `The temperature for the two beavers was highly correlated (`r apa_corr(beavers_corr, digits = 2)`). ``

#### Output

The temperature for the two beavers was highly correlated
(`r apa_corr(beavers_corr, digits = 2)`).

We can also format things like Bayes factors flexibly:

| Code | Output |
|------|--------|
| `format_bf(4321)`     |  `r format_bf(4321)`      |
| `format_bf(4321, digits1 = 2)`     | `r format_bf(4321, digits1 = 2)`       |
| `format_bf(4321, italics = FALSE, subscript = "")`     | `r format_bf(4321, italics = FALSE, subscript = "")`       |
| `format_bf(4321, cutoff = 1000)`     | `r format_bf(4321, cutoff = 1000)`       |
|      |        |
| `format_bf(0.04321)`     | `r format_bf(0.04321)`       |
| `format_bf(0.04321, digits2 = 3)`     | `r format_bf(0.04321, digits2 = 3)`       |
| `format_bf(0.04321, cutoff = 10)`     | `r format_bf(0.04321, cutoff = 10)`       |

## Citation

To cite formatstat, use:

> Stevens, Jeffrey R. (2023). formatstat: Format and print statistical output.
> (version 0.0.0.9000)
> <https://github.com/JeffreyRStevens/formatstat>
