---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# formatstats

<!-- badges: start -->

[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/JeffreyRStevens/formatstats/branch/main/graph/badge.svg)](https://app.codecov.io/gh/JeffreyRStevens/formatstats?branch=main)
<!-- badges: end -->

The goal of formatstats is to provide functions that flexibly format 
statistical output in a way that can be inserted into R Markdown documents. 
This is analogous to the
[`apa_print()`](https://frederikaust.com/papaja_man/reporting.html#statistical-models-and-tests)
functions in the [papaja](https://github.com/crsh/papaja) package, but 
functions in this package can print Markdown or LaTeX syntax. If your output 
document is a PDF, this doesn't matter. But if your output document is a Word
document (as required by many journal publishers), Markdown syntax generates
editable output instead of an image of output. 
The default style for statistical output follows [American
Psychological Association style](https://apastyle.apa.org/), but some
defaults can be over-ridden to flexibly format output.

## Installation

You can install the development version of formatstats from
[GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("remotes")
remotes::install_github("JeffreyRStevens/formatstats")
```

## Example

For an example, we'll create a correlation from the `mtcars` data set.

```{r example}
library(formatstats)
(cars_corr <- cor.test(mtcars$mpg, mtcars$disp))
```

Now we can apply the `format_corr()` function to `cars_corr` to create a
Markdown-formatted character string for the statistical results. We can
embed this as inline R Markdown code to generate the results.

#### Code

``Fuel efficiency and engine displacement were highly correlated (`r knitr::inline_expr('format_corr(cars_corr)')`). ``

#### Output

Fuel efficiency and engine displacement were highly correlated
(`r format_corr(cars_corr)`).

#### Other formatting

We can also format things like Bayes factors flexibly:

| Code | Output |
|------|--------|
| `format_bf(4321)`     |  `r format_bf(4321)`      |
| `format_bf(4321, digits1 = 2)`     | `r format_bf(4321, digits1 = 2)`       |
| `format_bf(4321, italics = FALSE, subscript = "")`     | `r format_bf(4321, italics = FALSE, subscript = "")`       |
| `format_bf(4321, cutoff = 1000)`     | `r format_bf(4321, cutoff = 1000)`       |
|      |        |
| `format_bf(0.04321)`     | `r format_bf(0.04321)`       |
| `format_bf(0.04321, digits2 = 3)`     | `r format_bf(0.04321, digits2 = 3)`       |
| `format_bf(0.04321, cutoff = 10)`     | `r format_bf(0.04321, cutoff = 10)`       |

## Citation

To cite formatstats, use:

> Stevens, Jeffrey R. (2023). formatstats: Format and print statistical output.
> (version 0.0.0.9000)
> <https://github.com/JeffreyRStevens/formatstats>


## Related packages

* [papaja](https://github.com/crsh/papaja)
* [apa](https://github.com/dgromer/apa)
