---
title: "formatstats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{formatstats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(formatstats)
```


## Formatting statistical objects

### Correlations


### T-tests


## Formatting numbers

### P-values


### Bayes factors

| Code | Output |
|------|--------|
| `format_bf(4321)`     |  `r format_bf(4321)`      |
| `format_bf(4321, digits1 = 2)`     | `r format_bf(4321, digits1 = 2)`       |
| `format_bf(4321, italics = FALSE, subscript = "")`     | `r format_bf(4321, italics = FALSE, subscript = "")`       |
| `format_bf(4321, cutoff = 1000)`     | `r format_bf(4321, cutoff = 1000)`       |
|      |        |
| `format_bf(0.04321)`     | `r format_bf(0.04321)`       |
| `format_bf(0.04321, digits2 = 3)`     | `r format_bf(0.04321, digits2 = 3)`       |
| `format_bf(0.04321, cutoff = 10)`     | `r format_bf(0.04321, cutoff = 10)`       |


### Scientific notation


### Other numbers



