---
title: "formatstats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{formatstats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(formatstats)
```

One of the most useful features of the [{papaja}](https://github.com/crsh/papaja) package is the [`apa_print()`](https://frederikaust.com/papaja_man/reporting.html#statistical-models-and-tests)
function, which takes a statistical object and formats the output to print the statistical information inline in R Markdown documents. The `apa_print()` function is an easy way to extract and format this statistical information for documents following APA style. However, APA style has some rather strange quirks, and users may want some flexibility in how their statistics are formatted. Moreover, `apa_print()` uses LaTeX syntax, which works great for PDFs but generates images for mathematical symbols when outputting to Word documents. 

The `{formatstats}` package uses APA style as the default, but allow more flexible formatting such as including the leading 0 before numbers with maximum values of 1. It also defaults to using Markdown syntax (instead of LaTeX) for easy conversions to Word documents.

## Formatting statistics

### Means and error

Often, we need to include simple descriptive statistics in our documents, such as measures of central tendency and error.

### P-values


### Bayes factors

| Code | Output |
|------|--------|
| `format_bf(4321)`     |  `r format_bf(4321)`      |
| `format_bf(4321, digits1 = 2)`     | `r format_bf(4321, digits1 = 2)`       |
| `format_bf(4321, italics = FALSE, subscript = "")`     | `r format_bf(4321, italics = FALSE, subscript = "")`       |
| `format_bf(4321, cutoff = 1000)`     | `r format_bf(4321, cutoff = 1000)`       |
|      |        |
| `format_bf(0.04321)`     | `r format_bf(0.04321)`       |
| `format_bf(0.04321, digits2 = 3)`     | `r format_bf(0.04321, digits2 = 3)`       |
| `format_bf(0.04321, cutoff = 10)`     | `r format_bf(0.04321, cutoff = 10)`       |


### Correlations

`r format_corr(cor.test(mtcars$mpg, mtcars$disp, method = "spearman"))`

`r format_corr(cor.test(mtcars$mpg, mtcars$disp, method = "spearman"), italics = FALSE)`

`r format_corr(cor.test(mtcars$mpg, mtcars$disp, method = "kendall"))`

`r format_corr(cor.test(mtcars$mpg, mtcars$disp, method = "kendall"), italics = F)`

### T-tests


## Formatting numbers

### Scientific notation


### Other numbers



